Concentration of Maxima and Fundamental Limits in High-Dimensional Testing and Inference

### 2021-02-01 3.phase_transitions/3.Gaussian-support-recovery.tex
merge of Ch. 3 and the new Ch. 4 & 5
I merged my work on Ch 3 and Ch. 4&5 with the "revision-GWAS" branch.

Need some more work -- putting the Mill's ratio in Ch 2 perhaps...

Turning to support recovery problems in the Gaussian error model \eqref{eq:model-additive-Chapter3},
we will analyze the asymptotic performance limits in terms of the risk metrics for exact, exact-approximate, approximate-exact support recovery problems (i.e., \eqref{eq:risk-exact}, \eqref{eq:risk-exact-approx}, and \eqref{eq:risk-approx-exact}, respectively), as well as the probability of support recovery \eqref{eq:risk-prob}. 
We will also review the recent result for exact support recovery risk \eqref{eq:risk-approximate} by \cite{arias2017distribution}, to reveal a rather complete landscape of support recovery problems in high-dimensional Gaussian error models.

We restrict our attention to the class of thresholding procedures in this section.
Specifically, the lower bounds that we develop in Theorems \ref{thm:Gaussian-error-exact-boundary} through \ref{thm:Gaussian-error-approx-exact-boundary} below are only meant to apply to thresholding procedures. 
Although it is intuitively appealing to consider only data-thresholding procedures in multiple testing problems, such procedures are not always optimal in more general settings. 
% Recently, \cite{arias2019detection} showed that thresholding procedures are in fact sub-optimal in the additive models \eqref{eq:model-additive} when errors have heavy (regularly-varying) tails. 
% Gao and Stoev [14] characterized the conditions under which thresholding procedures are optimal in the exact support recovery problem. 
% The optimality of data-thresholding procedures in terms of other statistical risks is an open problem that invites a dedicated investigation in a future study.
The optimality of thresholding procedures and the consequences of this restriction will be treated in Chapter \ref{chap:exact-support-recovery}.

A technical ingredient is needed in order to state our main results.
We define a rate at which the nominal levels of FWER or FDR go to zero.
\begin{definition} \label{def:slowly-vanishing}
We say the nominal level of errors $\alpha = \alpha_p$ vanishes slowly, if
\begin{equation} \label{eq:slowly-vanishing-error}
    \alpha\to 0,\quad \text{and} \quad \alpha p^\delta\to\infty \text{  for any } \delta>0.
\end{equation}
\end{definition}
As an example, the sequence of nominal levels $\alpha_p = 1/\log{(p)}$ is slowly vanishing, while the sequence $\alpha_p = 1/\sqrt{p}$ is not.

\subsection{The exact support recovery problem}
\label{subsec:exact-support-recovery-Gaussian}

@@

\begin{theorem} \label{thm:Gaussian-error-exact-boundary}
Consider the high-dimensional additive error model \eqref{eq:model-additive-Chapter3} under independent standard Gaussian errors, with signal sparsity and size as described in \eqref{eq:signal-sparsity-additive} and \eqref{eq:signal-size-additive}.
The function \eqref{eq:strong-classification-boundary-Gaussian} characterizes the phase transition of the exact support recovery problem.
Specifically, if $\underline{r} > {{g}}(\beta)$, then Bonferroni's, Sid\'ak's, Holm's, and Hochberg's procedures with slowly vanishing  nominal FWER levels (as defined in Definition \ref{def:slowly-vanishing}) all achieve asymptotically exact support recovery in the sense of \eqref{eq:support-recovery-success}.

Conversely, if $\overline{r} < {{g}}(\beta)$, then for any thresholding procedure $\widehat{S}_p$, we have $\P[\widehat{S}_p=S_p]\to0$.
Therefore, in view of Lemma \ref{lemma:risk-exact-recovery-probability}, exact support recovery asymptotically fails for all thresholding procedures in the sense of \eqref{eq:support-recovery-failure}.
\end{theorem}

We visualize the result in a $\beta$-$r$ phase diagram in Figure \ref{fig:phase-Gaussian-errors}.

Theorem \ref{thm:Gaussian-error-exact-boundary} is in fact a special case of the more general Theorem \ref{thm:sufficient} which covers dependent and non-Gaussian errors.
We will study the exact support recovery problem in greater detail, and prove the more general version of the Theorem in Chapter \ref{chap:exact-support-recovery}.

\begin{figure}
  \begin{center}
    \includegraphics[width=0.7\textwidth]{./figures/phase_diagram_Gaussian_ALL_boundaries.eps}
  \end{center}
   \caption{The phase diagram of support recovery problems for the high-dimensional chi-square model \eqref{eq:model-additive-Chapter3}, illustrating the boundaries of the exact support recovery (FWER + FWNR; top curve; Theorem \ref{thm:Gaussian-error-exact-boundary}), the approximate-exact support recovery (FDR + FWNR; second curve from top; Theorem \ref{thm:Gaussian-error-approx-exact-boundary}), the exact-approximate support recovery (FWER + FNR; horizontal line $r=1$; Theorem \ref{thm:Gaussian-error-exact-approx-boundary}), and the approximate support recovery problems (FDR + FNR; tilted line $r=\beta$; Theorem \ref{thm:Gaussian-error-approx-boundary}). The signal detection problem (Type I + Type II errors of the global test; lower curve) was studied in Donoho and Jin (2004). In each region of the diagram and above, the annotated statistical risk can be made to vanish, as dimension $p$ diverges. Conversely, the risks has liminf at least one.}
   \label{fig:phase-Gaussian-errors}
\end{figure}

\subsection{The approximate support recovery problem}
\label{subsec:approx-support-recovery-Gaussian}

@@

The analysis therein, however, assumed equal signal sizes for the alternatives.
We generalize the main results of \citet{arias2017distribution} to allow for unequal signal sizes.

\begin{theorem} \label{thm:Gaussian-error-approx-boundary}
In the context of Theorem \ref{thm:Gaussian-error-exact-boundary}, the function 
\begin{equation} \label{eq:approx-boundary-Gaussian}
    h(\beta) = \beta
\end{equation}
characterizes the phase transition of approximate support recovery problem.
Specifically, if $\underline{r} > {h}(\beta)$, then the Benjamini-Hochberg procedure (defined in Section \ref{sec:statistical-procedures}) with slowly vanishing nominal FDR levels (as defined in Definition \ref{def:slowly-vanishing}) achieves asymptotically approximate support recovery in the sense of \eqref{eq:support-recovery-success}.

Conversely, if $\overline{r} < {h}(\beta)$, then approximate support recovery asymptotically fails in the sense of \eqref{eq:support-recovery-failure} for all thresholding procedures.
\end{theorem}

Proof of Theorem \ref{thm:Gaussian-error-approx-boundary} is presented in Section \ref{sec:proofs}. 
The key to proving this generalization is a monotonicity property of the \ac{BH} procedure. 
Namely, the power of the \ac{BH} procedure in terms of FNR monotonically increases for stochastically larger alternatives.
This fact will be formalized in Lemma \ref{lemma:monotonicity-BH-procedure}, and may be of independent interest.

\subsection{The exact-approximate support recovery problem}
\label{subsec:exact-approx-support-recovery-Gaussian}

@@

\begin{theorem} \label{thm:Gaussian-error-exact-approx-boundary}
In the context of Theorem \ref{thm:Gaussian-error-exact-boundary}, the function 
\begin{equation} \label{eq:exact-approx-boundary-Gaussian}
    \widetilde{g}(\beta) = 1
\end{equation}
characterizes the phase transition of exact-approximate support recovery problem.
Specifically, if $\underline{r} > \widetilde{g}(\beta)$, then the procedures listed in Theorem \ref{thm:Gaussian-error-exact-boundary} with slowly vanishing nominal FWER levels (as defined in Definition \ref{def:slowly-vanishing}) achieve asymptotically exact-approximate support recovery in the sense of \eqref{eq:support-recovery-success}.

Conversely, if $\overline{r} < \widetilde{g}(\beta)$, then for any thresholding procedure $\widehat{S}$, the exact-approximate support recovery fails in the sense of \eqref{eq:support-recovery-failure}.
\end{theorem}

Theorem \ref{thm:Gaussian-error-exact-approx-boundary} is proved in Section \ref{sec:proofs}. 
The phase transition boundary \eqref{eq:exact-approx-boundary-Gaussian} is visualized in Figure \ref{fig:phase-Gaussian-errors}.

\begin{remark}
Boundary \eqref{eq:exact-approx-boundary-Gaussian} was briefly suggested by \citet{arias2017distribution}.
Unfortunately, it was falsely claimed that the boundary characterized the phase transition of the \emph{exact} support recovery problem, and the alleged proof was left as an ``exercise to the reader''.
This exercise was completed in Chapter \ref{chap:exact-support-recovery}, where the correct boundary \eqref{eq:exact-boundary-chisquared} was identified.

>>>>

Turning to support recovery problems in the Gaussian error model \eqref{eq:model-additive-Chapter3}, in the rest of this chapter
we will analyze the asymptotic performance limits in terms of the risk metrics for exact, exact-approximate, approximate-exact support recovery problems (i.e., \eqref{eq:risk-exact}, \eqref{eq:risk-exact-approx}, and \eqref{eq:risk-approx-exact}, respectively), as well as the probability of support recovery
\eqref{eq:risk-prob}.  We will also review the recent result for exact support recovery risk \eqref{eq:risk-approximate} by \cite{arias2017distribution}, to 
reveal a rather complete landscape of support recovery problems in high-dimensional Gaussian error models.

\begin{figure}
  \begin{center}
    \includegraphics[width=0.7\textwidth]{./figures/phase_diagram_Gaussian_ALL_boundaries.eps}
  \end{center}
   \caption{The phase diagram of support recovery problems for the high-dimensional model \eqref{eq:model-additive-Chapter3}, illustrating the boundaries of the exact support recovery (FWER + FWNR; top curve; Theorem \ref{thm:Gaussian-error-exact-boundary}), the approximate-exact support recovery (FDR + FWNR; second curve from top; Theorem \ref{thm:Gaussian-error-approx-exact-boundary}), the exact-approximate support recovery (FWER + FNR; horizontal line $r=1$; Theorem \ref{thm:Gaussian-error-exact-approx-boundary}), and the approximate support recovery problems (FDR + FNR; tilted line $r=\beta$; Theorem \ref{thm:Gaussian-error-approx-boundary}). The signal detection problem (Type I + Type II errors of the global test; lower curve) was studied in Donoho and Jin (2004). In each region of the diagram and above, the annotated statistical risk can be made to vanish, as dimension $p$ diverges. Conversely, the risks has liminf at least one.}
   \label{fig:phase-Gaussian-errors}
\end{figure}

In the rest of this chapter, we restrict our attention to the class of thresholding procedures. Specifically, the lower bounds that we develop in 
Theorems \ref{thm:Gaussian-error-exact-boundary} through \ref{thm:Gaussian-error-approx-exact-boundary} below are only meant to apply to thresholding
 procedures.  Although it is intuitively appealing to consider only data-thresholding procedures in multiple testing problems, such procedures are not always optimal in more general settings. 
% Recently, \cite{arias2019detection} showed that thresholding procedures are in fact sub-optimal in the additive models \eqref{eq:model-additive} when errors have heavy (regularly-varying) tails. 
% Gao and Stoev [14] characterized the conditions under which thresholding procedures are optimal in the exact support recovery problem. 
% The optimality of data-thresholding procedures in terms of other statistical risks is an open problem that invites a dedicated investigation in a future study.
The optimality of thresholding procedures and the consequences of this restriction will be treated in Chapter \ref{chap:optimality}.

\stilian{
Figure \ref{fig:phase-Gaussian-errors} illustrates the rich landscape of phase transitions in support recovery for the various choices
of statistical risk for the family of thresholding estimators, established in the following sections.}{\fbox{ added these words and moved the figure here}}
We end this brief overview with a technical notion needed in order to state our main results.
We define a rate at which the nominal levels of FWER or FDR go to zero.
\begin{definition} \label{def:slowly-vanishing}
We say the nominal level of errors $\alpha = \alpha_p$ vanishes slowly, if
\begin{equation} \label{eq:slowly-vanishing-error}
    \alpha\to 0,\quad \text{and} \quad \alpha p^\delta\to\infty \text{  for any } \delta>0.
\end{equation}
\end{definition}
As an example, the sequence of nominal levels $\alpha_p = 1/\log{(p)}$ is slowly vanishing, while the sequence $\alpha_p = 1/\sqrt{p}$ is not.

\section{The exact support recovery problem}
\label{subsec:exact-support-recovery-Gaussian}

@@

\begin{theorem} \label{thm:Gaussian-error-exact-boundary}
Consider the high-dimensional additive error model \eqref{eq:model-additive-Chapter3} under independent standard Gaussian errors, with signal sparsity and size as described in \eqref{eq:signal-sparsity-additive} and \eqref{eq:signal-size-additive}.
The function \eqref{eq:strong-classification-boundary-Gaussian} characterizes the phase transition of the exact support recovery problem.
Namely, the following two results hold.\\

{\rm (i)} If $\underline{r} > {{g}}(\beta)$, then Bonferroni's, Sid\'ak's, Holm's, and Hochberg's procedures with slowly vanishing  nominal FWER levels 
(as defined in Definition \ref{def:slowly-vanishing}) all achieve asymptotically exact support recovery in the sense of \eqref{eq:support-recovery-success}. \\

{\rm (ii)} Conversely, if $\overline{r} < {{g}}(\beta)$, then for any thresholding procedure $\widehat{S}_p$, we have $\P[\widehat{S}_p=S_p]\to0$.
Therefore, in view of Lemma \ref{lemma:risk-exact-recovery-probability}, exact support recovery asymptotically fails for all thresholding procedures in the sense of \eqref{eq:support-recovery-failure}.
\end{theorem}

We illustrate this result with a $\beta$-$r$ phase diagram in Figure \ref{fig:phase-Gaussian-errors}.  Theorem \ref{thm:Gaussian-error-exact-boundary} is in 
fact a special case of the more general Theorem \ref{thm:sufficient}, below, which covers dependent as well as Gaussain and non-Gaussian errors.  We will study the {\em exact support recovery} problem in greater detail and generality in Chapter \ref{chap:exact-support-recovery}.

\section{The approximate support recovery problem}
\label{subsec:approx-support-recovery-Gaussian}

@@

The analysis therein, however, assumed equal signal sizes for the alternatives.
We generalize the main results of \citet{arias2017distribution} to allow for unequal signal sizes. 
The key to establishing this generalization is a monotonicity property of the \ac{BH} procedure, presented in the following 
Section \ref{sec:BH-monotonicity}. Namely, the power of the \ac{BH} procedure in terms of FNR monotonically increases for stochastically 
larger alternatives. This fact will be formalized in Lemma \ref{lemma:monotonicity-BH-procedure}, and may be of independent interest.

\begin{theorem} \label{thm:Gaussian-error-approx-boundary}
In the context of Theorem \ref{thm:Gaussian-error-exact-boundary}, the function 
\begin{equation} \label{eq:approx-boundary-Gaussian}
    h(\beta) = \beta
\end{equation}
characterizes the phase transition of approximate support recovery problem.  Specifically the following two results hold.\\

{\rm (i)} If $\underline{r} > {h}(\beta)$, then the Benjamini-Hochberg procedure (defined in Section \ref{sec:statistical-procedures}) with slowly 
vanishing nominal FDR levels (as defined in Definition \ref{def:slowly-vanishing}) achieves asymptotically approximate support recovery in the sense of \eqref{eq:support-recovery-success}.\\

{\rm (ii)}
Conversely, if $\overline{r} < {h}(\beta)$, then approximate support recovery asymptotically fails in the sense of \eqref{eq:support-recovery-failure} for all thresholding procedures.
\end{theorem}

\begin{proof}[Proof of the necessary condition in Theorem \ref{thm:Gaussian-error-approx-boundary}]
We first show part {\rm (ii)}.  That is, when $\overline{r} < \beta$, no thresholding procedure is able to achieve approximate support recovery.
The arguments are similar to that in Theorem 1 of \cite{arias2017distribution}, although we allow for unequal signal sizes.

Denote the distributions of $\mathrm{N}(0,1)$, $\mathrm{N}(\underline{\Delta}, 1)$, and $\mathrm{N}((\overline{\Delta}, 1)$ as $F_0$, $F_{\underline{a}}$, and $F_{\overline{a}}$ respectively.

% We first show the necessary condition, i.e., when $\overline{r}<\beta$, approximate support recovery cannot be achieved with any thresholding procedure.
% In particular, we show that the liminf of the sum of FDP and NDP is at least 1.

Recall that thresholding procedures are of the form
$$
\widehat{S}_p = \left\{i\,|\,x(i) > t_p(x)\right\}.
$$
Denote $\widehat{S} := \left\{i\,|\,x(i) > t_p(x)\right\}$, and $\widehat{S}(u) := \left\{i\,|\,x(i) > u\right\}$.
For any threshold $u\ge t_p$ we must have $\widehat{S}(u)\subseteq\widehat{S}$, and hence
\begin{equation} \label{eq:approx-boundary-proof-FDP-Gaussian}
    \text{FDP} := \frac{|\widehat{S}\setminus{S}|}{|\widehat{S}|} \ge \frac{|\widehat{S}\setminus{S}|}{|\widehat{S}\cup{S}|} = \frac{|\widehat{S}\setminus{S}|}{|\widehat{S}\setminus{S}| + |S|} \ge
    \frac{|\widehat{S}(u)\setminus{S}|}{|\widehat{S}(u)\setminus{S}| + |S|}.
\end{equation}
On the other hand, for any threshold $u\le t_p$ we must have $\widehat{S}(u)\supseteq\widehat{S}$, and hence
\begin{equation} \label{eq:approx-boundary-proof-NDP-Gaussian}
    \text{NDP} := \frac{|{S}\setminus\widehat{S}|}{|{S}|} \ge 
    \frac{|{S}\setminus\widehat{S}(u)|}{|{S}|}.
\end{equation}
Since either $u\ge t_p$ or  $u\le t_p$ must take place, putting \eqref{eq:approx-boundary-proof-FDP-Gaussian} and \eqref{eq:approx-boundary-proof-NDP-Gaussian} together, we have
\begin{equation} \label{eq:approx-boundary-proof-converse-1-Gaussian}
    \text{FDP} + \text{NDP} 
    \ge \frac{|\widehat{S}(u)\setminus{S}|}{|\widehat{S}(u)\setminus{S}|+|{S}|} \wedge \frac{|{S}\setminus\widehat{S}(u)|}{|{S}|},
\end{equation}
for any $u$.
Therefore it suffices to show that for a suitable choice of $u$, the RHS of \eqref{eq:approx-boundary-proof-converse-1-Gaussian} converges to 1 in probability; the desired conclusion on FDR and FNR follows by the dominated convergence theorem.

Let $t^* = \sqrt{2q\log{p}}$ for some fixed $q$, we obtain an estimate of the tail probability by Mill's ratio \eqref{eq:Mills-ratio}, 
\begin{equation}
    \overline{F_0}(t^*) 
    \sim \frac{1}{t^*}\phi(t^*)
    = \frac{1}{2\sqrt{\pi q\log{p}}} p^{-q}, \label{eq:approx-boundary-proof-null-tail-prob-Gaussian}
\end{equation}
where $a_p\sim b_p$ is taken to mean $a_p/b_p\to 1$.
Observe that $|\widehat{S}(t^*)\setminus{S}|$ has distribution $\text{Binom}(p-s, \overline{F_0}(t^*))$ where $s=|S|$, denote $X = X_p := {|\widehat{S}(t^*)\setminus{S}|}/{|S|}$, and we have 
$$
\mu := \E\left[X\right] = \frac{(p-s)\overline{F_0}(t^*)}{s},
\quad \text{and} \quad
\var\left(X\right) = \frac{(p-s)\overline{F_0}(t^*){F_0}(t^*)}{s^2} \le \mu/s.
$$
Therefore for any $M>0$, we have, by Chebyshev's inequality,
\begin{equation}
    \P\left[X < M\right] 
    \le \P\left[\left|X-\mu\right| > \mu - M\right]
    \le \frac{\mu/s}{(\mu-M)^2}
    = \frac{1/(\mu s)}{(1-M/\mu)^2}. \label{eq:approx-boundary-proof-converse-2-Gaussian}
\end{equation}
Now, from the expression of $\overline{F_0}(t^*)$ in \eqref{eq:approx-boundary-proof-null-tail-prob-Gaussian}, we obtain
$$
\mu = (p^\beta - 1)\overline{F_0}(t^*) \sim \frac{1}{2\sqrt{\pi q\log{p}}} p^{\beta-q}.
$$
Since $\overline{r}<\beta$, we can pick $q$ such that $\overline{r}<q<\beta$. 
In turn, we have $\mu \to\infty$, as $p\to\infty$.
Therefore the last expression in \eqref{eq:approx-boundary-proof-converse-2-Gaussian} converges to 0, and we conclude that $X\to\infty$ in probability, and hence
\begin{equation} \label{eq:approx-boundary-proof-converse-3-Gaussian}
\frac{|\widehat{S}(t^*)\setminus{S}|}{|\widehat{S}(t^*)\setminus{S}|+|{S}|} 
= \frac{X}{X+1} \to 1 \quad \text{in probability}.
\end{equation}

On the other hand, we show that with the same choice of $u = t^*$, we have,
\begin{equation} \label{eq:approx-boundary-proof-converse-4-Gaussian}
    \frac{|{S}\setminus\widehat{S}(t^*)|}{|{S}|}\to 1 \quad \text{in probability}.
\end{equation}
By the stochastic monotonicity of Gaussian location family \eqref{eq:stochastic-monotonicity-Gaussian}, we have the following lower bound for the probability of missed detection for each signal $\mu(i)$, $i\in S$, 
\begin{equation} \label{eq:approx-boundary-proof-converse-5-Gaussian}
    \P[\mathrm{N}(\mu(i), 1) \le t^*] \ge F_{\overline{a}}(t^*).
\end{equation}
Since $|{S}\setminus\widehat{S}(t^*)|$ can be written as the sum of $s$ independent Bernoulli random variables,
\begin{equation*}
    |{S}\setminus\widehat{S}(t^*)| = \sum_{i\in S} \mathbbm{1}_{(-\infty, t^*]}(x(i)),
\end{equation*}
using with \eqref{eq:approx-boundary-proof-converse-5-Gaussian}, we conclude that $|{S}\setminus\widehat{S}(t^*)| \stackrel{\mathrm{d}}{\ge} \text{Binom}(s, {F_{\overline{a}}}(t^*))$.
Finally, we know that ${F_{\overline{a}}}(t^*)$ converges to 1 by our choice of diverging $t^*$, and the necessary condition is shown.
\end{proof}

\begin{proof}[Proof of the sufficient condition in Theorem \ref{thm:Gaussian-error-approx-boundary}]
We now turn to the sufficient condition, i.e., part {\rm (i)}. That is, when $\underline{r} > \beta$, the Benjamini-Hochberg procedure with 
slowly vanishing FDR levels achieves asymptotic approximate support recovery.

The FDR vanishes by our choice of $\alpha$ and the FDR-controlling property of the BH procedure \citep{benjamini1995controlling}.
It only remains to show that FNR also vanishes.

To do so we compare the FNR under the alternative specified in Theorem \ref{thm:Gaussian-error-approx-boundary} to one with all of the signal sizes equal to $\underline{\Delta}$.
By Lemma \ref{lemma:monotonicity-BH-procedure}, it suffices to show that the FNR under the BH procedure in this setting vanishes.
Let $x(i)$ be vectors of independent observations with $p-s$ nulls having standard Gaussian distributions, and $s$ signals having $\mathrm{N}(\underline{\Delta}, 1)$ distributions.

Denote the null and the alternative distributions as $F_0$ and $F_{a}$ respectively.
Let $\widehat{G}$ denote the empirical survival function as in \eqref{eq:empirical-tail-distribution}.
Define the empirical survival functions for the null part and signal part
\begin{equation} \label{eq:empirical-survival-null-signal-Gaussian}
    \widehat{W}_\text{null}(t) = \frac{1}{p-s}\sum_{i\not\in S}\mathbbm{1}\{x(i) \ge t\},
    \quad
    \widehat{W}_\text{signal}(t) = \frac{1}{s}\sum_{i\in S}\mathbbm{1}\{x(i) \ge t\},
\end{equation}
where $s=|S|$, so that
$$
\widehat{G}(t) = \frac{p-s}{p}\widehat{W}_\text{null}(t) + \frac{s}{p}\widehat{W}_\text{signal}(t).
$$

We need the following result to describe the deviations of the empirical distributions.
\begin{lemma}[Theorem 1 of \citet{eicker1979asymptotic}] \label{lemma:empirical-process}
Let $Z_1,\ldots,Z_k$ be iid with continuous survival function $Q$.
Let $\widehat{Q}_k$ denote their empirical survival function and define 
$\xi_k = \sqrt{2\log{\log{(k)}}/k}$ for $k \ge 3$. 
Then
$$
\frac{1}{\xi_k}\sup_z\frac{|\widehat{Q}_k(z) - Q(z)|}{\sqrt{Q(z)(1 - Q(z))}} \to 1,
$$
in probability as $k \to \infty$.
In particular,
$$
\widehat{Q}_k(z) = Q(z) + O_\P\left(\xi_k\sqrt{Q(z)(1 - Q(z))}\right),
$$
uniformly in z.
\end{lemma}

Apply Lemma \ref{lemma:empirical-process} to the two summands in $\widehat{G}$, we obtain
$\widehat{G}(t) = G(t) + \widehat{R}(t)$,
where 
\begin{equation} \label{eq:empirical-process-mean-Gaussian}
    G(t) = \frac{p-s}{p}\overline{F_0}(t) + \frac{s}{p}\overline{F_a}(t),
\end{equation}
and 
\begin{equation} \label{eq:empirical-process-residual-Gaussian}
    \widehat{R}(t) = O_\P\left(\xi_p\sqrt{\overline{F_0}(t)F_0(t)} + \frac{s}{p}\xi_s\sqrt{\overline{F_a}(t)F_a(t)}\right),
\end{equation}
uniformly in $t$.

Recall (see proof of Lemma \ref{lemma:monotonicity-BH-procedure}) that the BH procedure is the thresholding procedure with threshold set at 
\begin{equation} \label{eq:approx-boundary-proof-tau-Gaussian}
    \tau = \inf\{t\,|\,\overline{F_0}(t)\le\alpha\widehat{G}(t)\}. 
    %= \min\{t\,|\,\overline{F_0}(t)=\alpha\widehat{G}(t)\}.
\end{equation}
The NDP may also be re-written as 
$$
\text{NDP} = \frac{|{S}\setminus\widehat{S}|}{|{S}|} = \frac{1}{s}\sum_{i\in S}\mathbbm{1}\{x(i) < \tau\} = 1 - \widehat{W}_\text{signal}(\tau),
$$
so that it suffices to show that 
\begin{equation} \label{eq:approx-boundary-proof-sufficient-1-Gaussian}
    \widehat{W}_\text{signal}(\tau)\to 1
\end{equation} in probability.
Applying Lemma \ref{lemma:empirical-process} to $\widehat{W}_\text{signal}$, we know that 
$$
\widehat{W}_\text{signal}(\tau) = \overline{F_a}(\tau) + O_\P\left(\xi_s\sqrt{\overline{F_a}(\tau)F_a(\tau)}\right) = \overline{F_a}(\tau) + o_\P(1).
$$
So it suffices to show that $F_a(\tau)\to 0$ in probability.
Now let $t^* = \sqrt{2q\log(p)}$ for some $q$ such that $\beta<q<\underline{r}$.
We have 
\begin{equation} \label{eq:approx-boundary-proof-sufficient-2-Gaussian}
    F_a(t^*) 
    = \Phi(t^* - \underline{\Delta}) 
    = \Phi(\sqrt{2(q - \underline{r})\log{p}}) \to 0. 
\end{equation}
Hence in order to show \eqref{eq:approx-boundary-proof-sufficient-1-Gaussian}, it suffices to show 
\begin{equation} \label{eq:approx-boundary-proof-sufficient-3-Gaussian}
    \P\left[\tau \le t^*\right] \to 1.
\end{equation}
By \eqref{eq:empirical-process-mean-Gaussian}, the mean of the empirical process $\widehat{G}$ evaluated at $t^*$ is
\begin{equation} \label{eq:approx-boundary-proof-sufficient-4-Gaussian}
    G(t^*) = \frac{p-s}{p}\overline{F_0}(t^*) + \frac{s}{p}\overline{F_a}(t^*).
\end{equation}
The first term, using Relation \eqref{eq:approx-boundary-proof-null-tail-prob-Gaussian}, is asymptotic to $p^{-q}L(p)$, where $L(p)$ is the logarithmic term in $p$.
The second term, since $\overline{F_a}(t^*)\to 1$ by Relation \eqref{eq:approx-boundary-proof-sufficient-2-Gaussian}, is asymptotic to $p^{-\beta}$.
Therefore, $G(t^*) \sim p^{-q}L(p) + p^{-\beta} \sim p^{-\beta}$, since 
$p^{\beta-q}L(p)\to0$ where $q>\beta$.

The fluctuation of the empirical process at $t^*$, by Relation \eqref{eq:empirical-process-residual-Gaussian}, is 
\begin{align*}
    \widehat{R}(t^*) 
    &= O_\P\left(\xi_p\sqrt{\overline{F_0}(t^*)F_0(t^*)} + \frac{s}{p}\xi_s\sqrt{\overline{F_a}(t^*)F_a(t^*)}\right)\\
    &= O_\P\left(\xi_p\sqrt{\overline{F_0}(t^*)}\right) + o_\P\left(p^{-\beta}\right).
\end{align*}
By \eqref{eq:approx-boundary-proof-null-tail-prob-Gaussian} and the expression for $\xi_p$, the first term is $O_\P\left(p^{-(q+1)/2}L(p)\right)$ where $L(p)$ is a poly-logarithmic term in $p$.
Since $\beta<\min\{q,1\}$, we have $\beta<(q+1)/2$, and hence $\widehat{R}(t^*) = o_\P(p^{-\beta})$.

Putting the mean and the fluctuation of $\widehat{G}(t^*)$ together, we obtain
$$
\widehat{G}(t^*) = G(t^*) + \widehat{R}(t^*) \sim_\P G(t^*) \sim p^{-\beta},
$$
and therefore, together with \eqref{eq:approx-boundary-proof-null-tail-prob-Gaussian}, we have
$$
\overline{F_0}(t^*)/\widehat{G}(t^*) = p^{\beta-q}L(p)(1+o_{\P}(1)),
$$
which is eventually smaller than the FDR level $\alpha$ by the assumption \eqref{eq:slowly-vanishing-error} and the fact that $\beta<q$.
That is, 
$$
\P\left[\overline{F}_0(t^*) / \widehat{G}(t^*) < \alpha\right] \to 1.
$$
By definition of $\tau$ (recall \eqref{eq:approx-boundary-proof-tau-Gaussian}), this implies that $\tau \le t^*$ with probability tending to 1, and \eqref{eq:approx-boundary-proof-sufficient-3-Gaussian} is shown.
The proof for the sufficient condition is complete.
\end{proof}

\section{Monotonicity of the Benjamini-Hochberg procedure}
\label{sec:BH-monotonicity}
\input{3.phase_transitions/3.BH_monotonicity.tex}

\section{The exact-approximate support recovery problem}
\label{subsec:exact-approx-support-recovery-Gaussian}

@@

\begin{theorem} \label{thm:Gaussian-error-exact-approx-boundary}
In the context of Theorem \ref{thm:Gaussian-error-exact-boundary}, the function 
\begin{equation} \label{eq:exact-approx-boundary-Gaussian}
    \widetilde{g}(\beta) = 1
\end{equation}
characterizes the phase transition of exact-approximate support recovery problem.  Namely, the following two results hold.\\

{\rm (i)} If $\underline{r} > \widetilde{g}(\beta)$, then the procedures listed in Theorem \ref{thm:Gaussian-error-exact-boundary} with slowly vanishing nominal FWER levels (as defined in Definition \ref{def:slowly-vanishing}) achieve asymptotically exact-approximate support recovery in the sense of \eqref{eq:support-recovery-success}. \\

{\rm (ii)} Conversely, if $\overline{r} < \widetilde{g}(\beta)$, then for any thresholding procedure $\widehat{S}$, the exact-approximate support recovery fails in the sense of \eqref{eq:support-recovery-failure}.
\end{theorem}

The phase transition boundary \eqref{eq:exact-approx-boundary-Gaussian} is visualized in Figure \ref{fig:phase-Gaussian-errors}. The proof of this result
uses ideas from the proof of Theorem \ref{thm:Gaussian-error-approx-boundary} and is substantially shorter.

\begin{proof}[Theorem \ref{thm:Gaussian-error-exact-approx-boundary}]
We first show the sufficient condition. 
Vanishing FWER is guaranteed by the properties of the procedures, and we only need to show that FNR also goes to zero. 
Similar to the proof of Theorem \ref{thm:Gaussian-error-approx-boundary}, it suffices to show that
\begin{equation} \label{eq:additive-error-exact-approx-boundary-proof-sufficient-1}
    \text{NDP} = 1 - \widehat{W}_\text{signal}(t_p) \to 0,
\end{equation}
where $t_p$ is the threshold of Bonferroni's procedure.

Since $\alpha$ vanishes slowly (see Definition \ref{eq:slowly-vanishing-error}), for any $\delta>0$, we have $p^{-\delta}=o(\alpha)$.
Therefore, we have $-\log\alpha\le\delta\log{p}$ for large $p$, and
\begin{equation*} 
    1 \le \limsup_{p\to\infty}\frac{2\log{p} - 2\log{\alpha}}{2\log{p}} \le 1+\delta,
\end{equation*}
for any $\delta>0$.
Therefore, by the expression for normal quantiles, we know that 
$$
t_p=F^\leftarrow(1-\alpha/p)\sim(2\log{p}-2\log{\alpha})^{1/2} \sim(2\log{p})^{1/2}.
$$

Since $\underline{r}>\widetilde{g}(\beta)=1$, we can pick $q$ such that $1<q<\underline{r}$.
Let $t^* = \sqrt{2q\log{p}}$, we know that $t_p<t_p^*$ for large $p$.
Therefore for large $p$, we have
$$
\widehat{W}_\text{signal}(t_p) \ge \widehat{W}_\text{signal}(t^*) \ge \overline{F_a}(t^*) + o_\P(1),
$$
where $\overline{F_a}$ is the survival function of $\mathrm{N}(\sqrt{2\underline{r}\log{p}}, 1)$; the last inequality follows from the stochastic monotonicity of the Gaussian location family \eqref{eq:stochastic-monotonicity-Gaussian}, and Lemma \ref{lemma:empirical-process}.
Indeed, by our choice of $q<\underline{r}$, we obtain
$$
F_a(t^*) = \Phi\left(\sqrt{2(q-\underline{r})\log{p}}\right)\to0,
$$
and \eqref{eq:additive-error-exact-approx-boundary-proof-sufficient-1} is shown. 
This completes the proof of the sufficient condition.

The proof of the necessary condition follows similar structure as in the proof of Theorem \ref{thm:Gaussian-error-approx-boundary}, and uses the lower bound
\begin{equation} \label{eq:additive-error-exact-approx-boundary-proof-necessary-1}
    \mathrm{FWER}(\mathcal{R}) + \mathrm{FNR}(\mathcal{R}) \ge \P\left[\max_{i\in S^c}x(i)>u\right] \wedge \E\left[\frac{|S\setminus \widehat{S}(u)|}{|S|}\right],
\end{equation}
which holds for any arbitrary thresholding procedure $\mathcal{R}$ and arbitrary real $u\in\R$.

By the assumption that $\overline{r}<\widetilde{g}(\beta)=1$, we can pick $q$ such that $\overline{r}<q<1$ and let $u = t^*=\sqrt{2q\log{p}}$ in \eqref{eq:additive-error-exact-approx-boundary-proof-necessary-1}.
By relative stability of iid Gaussian random variables \eqref{eq:relative-stability-Gaussian-maxima}, we have
\begin{equation} \label{eq:additive-error-exact-approx-boundary-proof-necessary-2}
    \P\left[\frac{\max_{i\in S^c} x(i)}{\sqrt{2\log{p}}} > \frac{t^*}{\sqrt{2\log{p}}}\right] \to 1.
\end{equation}
since the first fraction in \eqref{eq:additive-error-exact-approx-boundary-proof-necessary-2} converges to 1, while the second converges to $q<1$.
Therefore, the first term on the right-hand side of \eqref{eq:additive-error-exact-approx-boundary-proof-necessary-1} converges to 1.
